---
title: "bits_and_pieces"
output: html_notebook
---

This is a file to collect lose code that I generate in the process of
making the final html files in my response to Michael's A-exam question.


# Folding Output

This is a stack overflow post about how to actually fold the output from an
R analysis:
- https://stackoverflow.com/questions/37755037/how-to-add-code-folding-to-output-chunks-in-rmarkdown-html-documents

# DATAOVERVIEW

I used the following code to do some cleaning for the statements csv, which
I don't actually present in the analysis. (I also did some manipulation by
hand in the csv, so things aren't complete.) (Also, file names may have
changed.)

```{r}
stats_raw %>%
  rename(per_id = `ï..per_id`) %>%
  select(-matches("^x$|^x\\.[0-9]+"))

# then, I used write.csv to write this back to "statements_final.csv"
```

# FEATURE EXTRACTION

Check some basics from feature extraction

```{r}
stats_raw
stats_clean
```

Unique word proportions

```{r}
stats_clean2 %>%
  select(statement,
         n_words,
         n_unique,
         n_uniq_prop)
```

