---
title: "Feature Extraction (Overview & Preliminaries)"
output:
  html_document:
    df_print: paged
    code_folding: show
---

[return to [overview page](./hld_OVERVIEW.html)]

# Overview

There is a long history of trying to find cues to deception (Bunn, 2012). 
For example, psychologists like Paul Ekman (2009) have tried to identify 
sublte nonverbal emotional and behavioral cues (e.g. facial "micro-expressions") 
which may reveal liars. Perhaps most notoriously, polygraph tests have been 
claimed to provide insight into whether a peson is lying by examining changes 
in various physiological states -- pulse rate, respiration, skin conductivity,
and so on (National Research Council, 2003; Polygraph, 2019). More recently, 
claims have emerged that functional magnetic resonance imaging can be used to 
detect lies, by examining differences in blood oxygen levels to various brain areas
(Simpson, 2008; Lie detection, 2019). Yet another method of analysis that falls
in this tradition is the analysis of text to suss out truths from lies (Newman, 
Pennebaker, Berry, & Richards, 2003; Pennebaker, Mehl, & Niederhoffer, 2003;
Pérez-Rosas & Mihalcea, 2015). In research of this type, the cues used to detect 
lies are features extracted from text -- for example: parts of speech, the 
sentiment of words, and the linguistic complexity of sentences. My analysis
will be of this sort, Thus, the next few sections will be devoted to extracting
various textual features that my prove useful for lie detection.

# Preliminaries

In this rest of this page, I will just get some basic prelimary 
housekeeping matters in order that will make the extraction of features 
in the following sections easier. (Feel free to skip over this, not much happens.)

## Preliminaries: Packages

Again, I will start by loading relevant packages.

```{r, message=FALSE, warning=FALSE}
library(tidyverse) # cleaning and visualization
library(quanteda) # text analysis
```

## Preliminaries: Data

Again, since this is a new analysis, I must load in the data that will be
analyzed. These will be exactly the raw statement previewed in the 
["Data Overview"](./hld_DATAOVERVIEW.html).

```{r, message=FALSE, warning=FALSE}
# First, load in the statements
stats_raw <- read.csv(file = "statements_final.csv")

```

## Preliminaries: Basic transformations and cleaning

First, we are going to convert all the statements to lower case (to ensure
that words like "I" and "i" are treated as the same word).

```{r}
# create new object to store cleaned data
stats_clean <- stats_raw

# Convert all text in statements to character
stats_clean$statement <- as.character(stats_clean$statement)

# Check that this was done correctly
mode(stats_clean$statement)

# Now convert all these statements to lower case
stats_clean$statement <- char_tolower(stats_clean$statement)

```

Now I will preview a random subset of statements, just to check that they look like
what I want.

```{r}
stats_clean %>%
  sample_n(5) %>%
  select(statement,
         everything())
```


## Save output

And now we can save the cleaned output (as an R data object), for future import
and use subsequent analyses.


```{r}
# save file
save(stats_clean,
     file = "stats_clean.Rda")

# now remove stats_clean from global environment (to avoid later confusion)
rm(stats_clean)

```

# Citations

* Bunn, G. C. (2012). The truth machine: A social history of the lie detector. 
Johns Hopkins University Press.

* Ekman, P. (2009). Telling lies: Clues to deceit in the marketplace,
politics, and marriage (revised edition). WW Norton & Company.

* Lie detection. (2019). In Wikipedia. Retrieved from 
https://en.wikipedia.org/w/index.php?title=Lie_detection&oldid=887268550
 
* National Research Council. (2003). The polygraph and lie detection. 
National Academies Press.

* Newman, M. L., Pennebaker, J. W., Berry, D. S., & Richards, J. M. (2003).
Lying words: Predicting deception from linguistic styles. Personality and 
social psychology bulletin, 29(5), 665-675.

* Pennebaker, J. W., Mehl, M. R., & Niederhoffer, K. G. (2003). 
Psychological aspects of natural language use: Our words, our selves.
Annual Review of Psychology, 54(1), 547-577.

* Pérez-Rosas, V., & Mihalcea, R. (2015). Experiments in open domain deception 
detection. Proceedings of the 2015 Conference on Empirical Methods in Natural 
Language Processing, 1120-1125.

* Polygraph. (2019). In Wikipedia. Retrieved from
https://en.wikipedia.org/w/index.php?title=Polygraph&oldid=895112117

* Simpson, J. R. (2008). Functional MRI lie detection: too good to be true?
Journal of the American Academy of Psychiatry and the Law Online, 36(4), 491-498.

# END
