---
title: "Data Overview"
output:
  html_document:
    df_print: paged
    code_folding: show
---

[return to [overview page](./hld_OVERVIEW.html)]

Before delving deeper into the data, I will first give an overview of the data,
so that you might get a better "feel" for the basic structure and nature of the
responses (statements) that people provided. I will do this by going through each
of the six questions to which people responded, and showing a sample set of
responses for each question.

# Packages

Before loading in any of the data, I will first load some very useful packages
in R that will faciliated analysis and cleaning of the data.

```{r, message=FALSE, warning=FALSE}
# this is actually a series of packages, highly useful for cleaning and visualizing data
library(tidyverse)
```

# Import

Next, I will import the statements, which are stored in a csv file. I will
import and then store these statements in an object called "stats_raw" (i.e.
raw statements). (Note I have already done some pre-procesing, which I decided to
spare you from here. For example, I already eliminated responses from anyone
who did not respond to all six questions with both a true statement and a
false statement. I also eliminated some empty columns and renamed
columns to make their meaning clearer.)

[Note, that you can toggle the display of the code, by clicking the button 
labeled "hide"/"code" on the right, above each code block,]

```{r, message=FALSE, warning=FALSE}
# First, load in the statements
stats_raw <- read.csv(file = "statements_final.csv")

```

# Basics

First, let's look at some very basic features of the data set, to get a 
better understanding of its basic structure and organization.

To begin, let's look at the basic dimensions of the dataset -- that is the
number of rows and columns.

```{r}
dim(stats_raw)
```

Above, we see that the dataset has 5004 rows, and 13 columnns.

Let's examine what those columns are.

```{r}
# See the names of the columns
colnames(stats_raw)

```

As we see, we have the following columns

Column Name  | Content
------------- | -----------------------------------------------------------------
per_id  | unique identification number for each participant
stat_id  | unique identification number for each statement
q_num | the question (of the six) to which this statement is a response
grd_truth | "ground truth", a column indicating whether the statment is a lie or a truth
statement | this contains the entire text of the participants response
stat_method | for each lie, the participants written explanation for how they generated that lie (empty for truths)
stat_type | for each lie, the participant's multiple choice response for whether the response is based more on fantasy or reality (empty for truth)
order_first | the order in which participants completed the statements (lie followed by truths, or truths followed by lies)
sex | the participants self-reported sex
age | the participants self-reported age in years
race | the participants self-reported race/ethnicity
rand | a random number (previously used to sort the statements)
rand_order | a number ranking each statement by the previous random number (i.e. lowest random # gets 1, second lowest gets 2, etc)

Finally, let's take a little preview at a few full rows of the data, to also
get a better sense of the structure. (Note: the columns appear in the order
"per_id", "stat_id", "q_num", etc. You can view more columns, by clicking
the right pointing arrow following at the right of the right-most
column name.)

```{r}
head(stats_raw)
```


# Statements

Okay, now let's look at some of the responses provided by participants, to each
of the six questions.

## Question 1: Meeting

The first question instructed participants to "Describe how you met [person you know well]."
Let's look at some of the responses participants gave.

### Truthful Responses (Q1)

Here is a sample of 5 truthful responses.
(Note: I've re-arranged the columns here, so that the statement text appears as the
firsct column.)

```{r}
# display 5 true responses to question 1, by pre-computed random number
stats_raw %>%
  filter(q_num == 1,
         grd_truth == "truth") %>%
  arrange(rand_order) %>%
  select(statement,
         everything()) %>%
  top_n(5)

```

### Untruthful Responses, i.e. lies (Q1)

And here is a sample of 5 lies.

```{r}
# display 5 untruthful responses to question 1, by pre-computed random number
stats_raw %>%
  filter(q_num == 1,
         grd_truth == "lie") %>%
  arrange(rand_order) %>%
  select(statement,
         everything()) %>%
  top_n(5)

```


## Question 2: Regret

The second question asked participants: "What is something in your life that
you regret (i.e. wish you had done differently)?"

### Truthful Responses (Q2)

Here is a sample of 5 truthful responses.

```{r}
# display 5 true responses to question 2, by pre-computed random number
stats_raw %>%
  filter(q_num == 2,
         grd_truth == "truth") %>%
  arrange(rand_order) %>%
  select(statement,
         everything()) %>%
  top_n(5)

```

### Untruthful Responses, i.e. lies (Q2)

And here is a sample of 5 lies.

```{r}
# display 5 untruthful responses to question 2, by pre-computed random number
stats_raw %>%
  filter(q_num == 2,
         grd_truth == "lie") %>%
  arrange(rand_order) %>%
  select(statement,
         everything()) %>%
  top_n(5)

```


## Question 3: Yesterday

The third question instructed participants to: "Please describe what you did
yesterday."

### Truthful Responses (Q3)

Here is a sample of 5 truthful responses.

```{r}
# display 5 true responses to question 3, by pre-computed random number
stats_raw %>%
  filter(q_num == 3,
         grd_truth == "truth") %>%
  arrange(rand_order) %>%
  select(statement,
         everything()) %>%
  top_n(5)

```

### Untruthful Responses, i.e. lies (Q3)

And here is a sample of 5 lies.

```{r}
# display 5 untruthful responses to question 3, by pre-computed random number
stats_raw %>%
  filter(q_num == 3,
         grd_truth == "lie") %>%
  arrange(rand_order) %>%
  select(statement,
         everything()) %>%
  top_n(5)

```

## Question 4: Liking

The fourth question instructed participants to: "Give some reason why you
like [person you like/dislike]" (where the name of a person they listed as
as person they like was piped in when they were asked for a truthful response,
and the name of a person they listed as a person they disliked was piped in
when they were asked for a un untruthful response).

### Truthful Responses (Q4)

Here is a sample of 5 truthful responses.

```{r}
# display 5 true responses to question 4, by pre-computed random number
stats_raw %>%
  filter(q_num == 4,
         grd_truth == "truth") %>%
  arrange(rand_order) %>%
  select(statement,
         everything()) %>%
  top_n(5)

```

### Untruthful Responses, i.e. lies (Q4)

And here is a sample of 5 lies.

```{r}
# display 5 untruthful responses to question 4, by pre-computed random number
stats_raw %>%
  filter(q_num == 4,
         grd_truth == "lie") %>%
  arrange(rand_order) %>%
  select(statement,
         everything()) %>%
  top_n(5)

```

## Question 5: Strength

The fifth question instructed participants: "Describe a personal strength 
of yours (i.e. something that you are good at)."

### Truthful Responses (Q5)

Here is a sample of 5 truthful responses.

```{r}
# display 5 true responses to question 5, by pre-computed random number
stats_raw %>%
  filter(q_num == 5,
         grd_truth == "truth") %>%
  arrange(rand_order) %>%
  select(statement,
         everything()) %>%
  top_n(5)

```

### Untruth Responses, i.e. lies (Q5)

And here is a sample of 5 lies.

```{r}
# display 5 untruthful responses to question 5, by pre-computed random number
stats_raw %>%
  filter(q_num == 5,
         grd_truth == "lie") %>%
  arrange(rand_order) %>%
  select(statement,
         everything()) %>%
  top_n(5)

```

## Question 6: Hobby

In the sixth question, participants were asked: "What is a hobby of yours
or something that you enjoy doing in your free time? And why do you enjoy it?"

### Truthful Responses (Q6)

Here is a sample of 5 truthful responses.

```{r}
# display 5 true responses to question 6, by pre-computed random number
stats_raw %>%
  filter(q_num == 6,
         grd_truth == "truth") %>%
  arrange(rand_order) %>%
  select(statement,
         everything()) %>%
  top_n(5)

```


### Untruth Responses, i.e. lies (Q6)

And here is a sample of 5 lies.

```{r}
# display 5 untruthful responses to question 6, by pre-computed random number
stats_raw %>%
  filter(q_num == 6,
         grd_truth == "lie") %>%
  arrange(rand_order) %>%
  select(statement,
         everything()) %>%
  top_n(5)

```


# END