---
title: "Data Cleaning & Pre-Processing (continued)"
output:
  html_document:
    df_print: paged
    code_folding: show
---

[return to [overview page](./hld_OVERVIEW.html)]

Before beginning attempting to build models to predict the truth
or falsity of statements, we need to do some house-keeping. We need to join together,
examine, pre-process and clean the sets of features we created in the earlier sections.
That is what I do in this section.

# Packages

Again, I will start by loading relevant packages.

```{r, message=FALSE, warning=FALSE}
# before knitting: message = FALSE, warning = FALSE
library(tidyverse) # cleaning and visualization
library(plyr) # has join_all function
library(ggthemes) # visualization
library(caret) # modeling
library(AppliedPredictiveModeling)
library(e1071) # has skewness() function
library(DescTools) # has Winsorize() function
library(ggcorrplot) # for correlation plot
```

# Load Data

First, I will load the various data from the features we just extracted.

```{r}
# load all the nice tidy df's of features we created (remember stats_words has multiple dtm's)
load("stats_clean.Rda") # has text of statement
load("stats_length.Rda") # statement lengths
load("stats_pos.Rda") # parts of speech
load("stats_sent.Rda") # sentiment
load("stats_complex.Rda") # complexity and readability
load("stats_words.Rda") # bag of words (mini document-term matrices)

```
