---
title: "Hybrid Modeling (Support Vector Machine)"
author: Sebastian Deri
output:
  html_document:
    df_print: paged
    code_folding: show
---

[return to [overview page](./hld_OVERVIEW.html)]

In this section, we will generate our second hybrid human-computer model.
This will be a support vector machine model. The general training-testing and
model building procedure will be exactly the same as for our earlier
[support vector machine model](./hld_MODEL_logistic.html), with the exception
of course that our hybrid model will also take into account human predictions
as a feature (in addition to the textual features used earlier). The performance
of this model will be compared to the performance of a non-hybrid model (which uses
only textual features) trained on the same statements.

# Packages

Again, I will start by loading relevant packages.

```{r, message=FALSE, warning=FALSE}
# before knitting: message = FALSE, warning = FALSE
library(tidyverse) # cleaning and visualization
library(caret) # modeling
library(kernlab) # has sigest, for selecting tuning parameter sigma

```

# Load Data

Next, I will load the data file, we created [earlier](./hld_HMODEL_overview.html),
which has both the cleaned and processed textual features and human predictions.
Note, again, that these data files consists of a total of 3,663 statments.

```{r}
# load df of combined human and processed textual feature and ground truth
load("stats_combo.Rda")

# For rendering, I'm going to cheat here and load results created when this model was first run
# For some reason, chunks that were supposed to be cached when originally run are rerunning
# load("results_HYB_svm.Rda")
# change the specific names (renamed at end), back to generic name
# results_HYB_svm -> results
# model_hybrid_svm -> model_hybrid
# model_non_hybrid_svm -> model_non_hybrid

```

```{r}
# convert predict column from factor (truth, lie) to numeric (1, 0); for use by sigest
stats_combo <-
  stats_combo %>%
  mutate(predict = case_when(predict == "truth" ~ 1,
                             predict == "lie" ~ 0))

# print
stats_combo
```

# EXAMPLE (Single Support Vector Machine Model, with Radial Basis Function)

As usual, let's begin with a singular example. Here we will simply train and test
two support vector machine models: one hybrid model that takes into accounts
both human predictions and textual features and one that only takes into
account textual features. As with our original support vector machine models,
we will use radial basis kernel functions, to take advantage of SVMs ability 
to produce non-linear classification boundaries (see discussion in final
introductory paragraphs of earlier 
[SVM page, right above "Packages" section](./hld_MODEL_logistic.html),
for some more discussion).

## Split Sample Into Training and Testing Sets

As before, we are going to conduct an exact 50-50 split,
randomly allocating one half of the statements to the training set, and
the other half to the testing set (using the createDataPartition function in the 
caret packages; Kuhn, 2008; Kuhn, 2019). The exact same training set will be
used to build the hybrid and non-hybrid models, both of which will then also be tested
on the exact same testing set.

```{r, cache = FALSE}
# set seed, so that statistics don't keep changing for every analysis
set.seed(2019)

# partition data in 50-50 lgocv split (create index for test set)
index_train_ex <- 
  createDataPartition(y = stats_combo$stat_id,
                      p = 0.50,
                      times = 1,
                      list = FALSE)

# actually create data frame with training set (predictors and outcome together)
train_set_ex <- stats_combo[index_train_ex, ]

# actualy create data frame with test set (predictors and outcome together)
test_set_ex <- stats_combo[-index_train_ex, ]

```

## Build Model (on Training Set)

Now that the data are split, we can fit support vector machine models to the training
data. Again, we will fit two support vector machine models. One, a hybrid model that uses
both textual features and human guesses as predictors. An the other, a non-hybrid
model that uses on textual features as predictors. 

As with our earlier support vector machine models, and unlike with our logistic
regression models, this model has two tuning parameters that we need to set:
our cost penalty and our sigma parameter in the radial basis function. As before,
I will set the sigma parameter using the slightly modified recommendations from the
sigest function in the kernlab package. And as before, I will select a cost penalty
by cycling through the candidate values of 0.25, 0.5, 1, 2, and 4, and selecting
the value that leads to the highest level of overall accuracy across 3 repeated
50-50 training testing splits (see "Build Model (on Training Set)" section of earlier
[SVM page](./hld_MODEL_logistic.html) for more detailed discussion).


```{r svm_tune_params}
# note: these setting chunks are separated for reuse later

# set seed, so that statistics don't keep changing for every analysis
set.seed(2019)

# -----------------------------------------------------------------------------
# STEP 1: SELECT TUNING PARAMETERS

# part a: select cost penalty values
costs_svm <- c(0.25, 0.5, 1, 2, 4)

# part b: get suggested sigma value, for radial basis function, for this data set
# i take the median of the kernlab suggestions 
# (default in caret is to take mean, excluding second suggestion)
# see: https://github.com/topepo/caret/blob/master/models/files/svmRadial.R
# and: https://stats.stackexchange.com/questions/408159/what-is-the-basis-for-the-default-sigma-value-used-by-svmradial-in-caret
sigma_hyb_svm <- median(
              kernlab::sigest(
                as.matrix(train_set_ex 
                            %>% select(
                              -stat_id,
                              -grd_truth)
                          ),
                scaled = TRUE
                )
              )

sigma_non_hyb_svm <- median(
              kernlab::sigest(
                as.matrix(train_set_ex 
                            %>% select(
                              -stat_id,
                              -grd_truth,
                              -predict)
                          ),
                scaled = TRUE
                )
              )

# part c: save parameters in tune grid object
tune_grid_hyb_svm <- expand.grid(sigma = sigma_hyb_svm,
                                 C = costs_svm)

tune_grid_non_hyb_svm <- expand.grid(sigma = sigma_non_hyb_svm,
                                     C = costs_svm)

# -----------------------------------------------------------------------------
# STEP 2: SELECT TUNING METHOD
# set up train control object, which specifies training/testing technique
train_control_svm <- trainControl(method = "LGOCV",
                                  number = 3,
                                  p = 0.50)

```

```{r make_example_models, cache = TRUE, eval = TRUE}
# set seed, so that statistics don't keep changing for every analysis
# (applies for models which might have random parameters)
set.seed(2019)

# start timer
start_time <- Sys.time()

# -----------------------------------------------------------------------------
# STEP 3: TRAIN MODEL

# use caret "train" function to train svm
model_hybrid <- 
  train(form = grd_truth ~ . -stat_id,
        data = train_set_ex,
        method = "svmRadial",
        tuneGrid = tune_grid_hyb_svm,
        trControl = train_control_svm,
        metric = "Accuracy") # how to select among models

model_non_hybrid <- 
  train(form = grd_truth ~ . -stat_id -predict,
        data = train_set_ex,
        method = "svmRadial",
        tuneGrid = tune_grid_non_hyb_svm,
        trControl = train_control_svm,
        metric = "Accuracy") # how to select among models

# end timer
total_time <- Sys.time() - start_time

```

```{r}
# examine tuning parameters selected for non-hybrid model
model_non_hybrid

```


```{r}
# examine tuning parameters select for hybrid model
model_hybrid

```

## Evaluate Model (on Testing Set)

Finally, let's examine and compare the performance of the two models. As before,
this is done by using the models to generate predictions on the remaining ~50%
of the data (1831 statements) in the training set. And then examining the performance
statistics using the confusionMatrix function from the caret package (Kuhn, 2008).

Both models have an overall accuracy level that is above chance:

* hybrid model: 64.1% [95% CI: 61.9, 66.3%]
* non-hybrid model: 63.2% [95% CI: 60.9, 65.4%]

And the overall accuracy of the hybrid model is better than the overall accuracy
of the non-hybrid model. Dissapointingly, however, this difference is not 
significant, chi-squared = 0.30, p = 0.583. Not even close in fact. This is both
dissapointing and surprising. [Earlier](./hld_HUMAN_perf.html), we saw that human 
predictions alone achieved an accuracy of over 55%, which was well above chance.
And a model trained only on textual features here does well above chance as well
(overall accuracy > 63%, as seen just above). It would stand to reason that adding
in human predictions would serve to significantly improve the model. This is of
course not what happened. I will need to conduct further analyses to examine why this
is. One possible hypothesis at the moment is that (1) human predictions are 
somehow redundant with some existing feature or set of features. (Note: this is not to
say that such a feature or set of features is the actual basis of human judgments,
just that there might be, for whatever reason, an assocation). However, a priori,
this seems unlikely to me, knowing the features we have extracted (i.e. it seems
unlikely to me that human judgments are going to track something as clinical as
the number of pronouns in a sentence, or even more "human" features like sentiment).
Another hypothesis relates to (2) the different pattern of guessing by which
humans and computer models achieve their accuracy. As we saw in our previous 
analyses, computer models had roughly comparable levels of sensitivity (rate of 
truth detection) and specificity (rate of lie detection). However, humans had
much higher sensitivity than specificity. I am not sure of the exact effect of
this on the hybrid models, but a loose speculation might be that any gains in
accuracy that might come from adding in human predictions would come from
correctly identifying a larger portion of truths as truths, however, these would
be "offset" by increasing the portion of lies that are incorrectly identified.

With regard to other performance metrics, I have not yet conducted formal statistical
for each of theme, but sensitivity, specificity, precision, and negative predictive
value look roughly the same between the hybrid and non-hyrbid models.

```{r, cache = FALSE}
# generate predictions for hybrid model 
preds_hybrid <-
  predict(object = model_hybrid,
                 newdata = test_set_ex,
                 type = "raw")

# generate predictions for non-hybrid model
preds_non_hybrid <-
  predict(object = model_non_hybrid,
                 newdata = test_set_ex,
                 type = "raw")

```

```{r}
# examine hybrid model performance
conf_hybrid <-
  confusionMatrix(data = preds_hybrid,
                  reference = test_set_ex$grd_truth,
                  positive = "truth")

# print
conf_hybrid

```

```{r}
# examine non-hybrid model performance
conf_non_hybrid <-
  confusionMatrix(data = preds_non_hybrid,
                  reference = test_set_ex$grd_truth,
                  positive = "truth")

# print
conf_non_hybrid

```

```{r}
# calculate values relevant for test (and some extras, in case i want to compare sensitivity, etc)
hybrid_truth_correct = conf_hybrid$table[4]
hybrid_truth_total = conf_hybrid$table[3] + conf_hybrid$table[4]
hybrid_lie_correct = conf_hybrid$table[1]
hybrid_lie_total = conf_hybrid$table[1] + conf_hybrid$table[2]
hybrid_total_correct = hybrid_truth_correct + hybrid_lie_correct
hyrbid_total_truth_guesses = conf_hybrid$table[2] + conf_hybrid$table[4]
hyrbid_total_lie_guesses = conf_hybrid$table[1] + conf_hybrid$table[3]
hybrid_total = sum(conf_hybrid$table)

non_hybrid_truth_correct = conf_non_hybrid$table[4]
non_hybrid_truth_total = conf_non_hybrid$table[3] + conf_non_hybrid$table[4]
non_hybrid_lie_correct = conf_non_hybrid$table[1]
non_hybrid_lie_total = conf_non_hybrid$table[1] + conf_non_hybrid$table[2]
non_hybrid_total_correct = non_hybrid_truth_correct + non_hybrid_lie_correct
non_Hyrbid_total_truth_guesses = conf_non_hybrid$table[2] + conf_non_hybrid$table[4]
non_hyrbid_total_lie_guesses = conf_non_hybrid$table[1] + conf_non_hybrid$table[3]
non_hybrid_total = sum(conf_non_hybrid$table)

```


```{r}
# conduct actual test
prop.test(x = c(hybrid_total_correct, non_hybrid_total_correct),
          n = c(hybrid_total, non_hybrid_total),
          alternative = "two.sided",
          conf.level = 0.95)
```

```{r}
# examine z value, using two proportion z-test just for comparison

# function to compute z-value
# from: https://www.r-bloggers.com/comparison-of-two-proportions-parametric-z-test-and-non-parametric-chi-squared-methods/
z.prop = function(x1,x2,n1,n2){
  numerator = (x1/n1) - (x2/n2)
  p.common = (x1+x2) / (n1+n2)
  denominator = sqrt(p.common * (1-p.common) * (1/n1 + 1/n2))
  z.prop.ris = numerator / denominator
  return(z.prop.ris)
}

# actual z prop (z = 0.58)
z_value <-
  z.prop(x1 = hybrid_total_correct,
         x2 = non_hybrid_total_correct,
         n1 = hybrid_total, 
         n2 = non_hybrid_total)

# get corresponding p-value (two-sided) (p = 0.559)
p_value <- 2*pnorm(-abs(z_value))

# print z and p
print(paste("z=", round(z_value, 3),
            ", p=", round(p_value, 3),
            sep = ""))
```


# Change in predictions

Let's try to get some insight into how the incorporating human predictions
changed the predictions made by the hybrid model (as compared to the predictions
made by the non-hybrid model). As we have discussed [before](./hld_MODEL_overview.html),
there are only four possibile outcomes when the predictions of a model are compared
to reality: true positive, true negative, false positive, false negative.
All of the model's performance depends on these four quantities. We can
think of the non-hybrid model as a "baseline", upon which the hybrid model might
improve. The only way the hybrid model might improve upon the non-hybrid model
is through one of two ways: a statment that was incorrectly predicted as a false negative
by the non-hybrid model is correctly predicted as a true positive by the hybrid model,
or a statement that was incorrectly predicted as a false positive by the non-hybrid
model is correctly predicted as a tur negative by the hybrid model. The rest of the 
changes either have no effect (a true positive stays a true positive from the non-hybrid
to a hybrid model, a true negative stays a true negative, a false positive stays a 
false positive, or a false negative stays a false negative) or are detrimental 
(a true positive becomes a false negative, or a true negative becomes a false
positive).

Below, I have tabulated the total number of times these eight possible changes
have occurred, when the non-hybrid model's predictions are compared to the
hybrid model's predictions. What we see is that most of the predictions stayed
the same
between the non-hybrid and hybrid model (1698 of 1831 predictions, 92.7%, were the
same). And only a slightly higher number of predictions went from incorrect 
to correct (75 cases, 4.1%, were either false negative that become true positives
or false positives that become true negatives) than the number of predictions
that went from correct to incorrect (58 cases, 3.2%, were either true positives
that become false negatives or true negatives that became false positives).

More detailed analyses should examine how the human predictions relate to the
other features used for prediction.

```{r}
# combine: grd_truth, non_hybrid predictions, hybrd predictions
preds_comp <-
  cbind(as.data.frame(test_set_ex$grd_truth),
        as.data.frame(preds_non_hybrid),
        as.data.frame(preds_hybrid))

# label columns
colnames(preds_comp) <- c("grd_truth", "non_hyb", "hyb")

# print
# preds_comp

# generate additional information
preds_tally <-
  preds_comp %>%
  group_by(grd_truth, non_hyb, hyb) %>%
  summarize(n = n()) %>%
  arrange(desc(grd_truth), desc(non_hyb), desc(hyb)) %>%
  mutate(change_present = case_when(non_hyb == hyb ~ "no",
                            non_hyb != hyb ~ "yes")) %>%
  mutate(change_type = case_when(grd_truth == "truth" & non_hyb == "truth" & hyb == "truth" ~ "TP -> TP",
                            grd_truth == "truth" & non_hyb == "truth" & hyb == "lie" ~ "TP -> FN",
                            grd_truth == "truth" & non_hyb == "lie" & hyb == "truth" ~ "FN -> TP",
                            grd_truth == "truth" & non_hyb == "lie" & hyb == "lie" ~ "FN -> FN",
                            grd_truth == "lie" & non_hyb == "truth" & hyb == "truth" ~ "FP -> FP",
                            grd_truth == "lie" & non_hyb == "truth" & hyb == "lie" ~ "FP -> TN",
                            grd_truth == "lie" & non_hyb == "lie" & hyb == "truth" ~ "TN -> FP",
                            grd_truth == "lie" & non_hyb == "lie" & hyb == "lie" ~ "TN -> TN")) %>%
  mutate(change_effect = case_when(change_type == "TP -> TP" ~ "nothing",
                                   change_type == "TN -> TN" ~ "nothing",
                                   change_type == "FP -> FP" ~ "nothing",
                                   change_type == "FN -> FN" ~ "nothing",
                                   change_type == "FN -> TP" ~ "gain",
                                   change_type == "FP -> TN" ~ "gain",
                                   change_type == "TP -> FN" ~ "loss",
                                   change_type == "TN -> FP" ~ "loss")) %>%
  arrange(change_effect)

# print
preds_tally

```

```{r}
# tally the total number of times each type of "change_effect" occurred
preds_tally %>%
  group_by(change_effect) %>%
  summarize(n = sum(n))

```


# FULL (Hybrid and Non-Hybrid Logistic Regression Models)

To get more reliable estimates of the performance of non-hybrid and hybrid models,
let's repeat the exact model training and testing procedure from above 10 times,
storing and aggregating the results.

## Run 10 models

Below is the code that runs through this modeling process 10 different times and
saves the result from each round.

```{r make_FULL_models, cache = TRUE, eval = FALSE}
# # -----------------------------------------------------------------------------
# STEP 0: set seed, so that statistics don't keep changing for every analysis
set.seed(2019)

# # -----------------------------------------------------------------------------
# STEP 1: decide how many times to run the model
rounds <- 10

# -----------------------------------------------------------------------------
# STEP 2: set up object to store results
# part a: create names of results to store
result_cols <- c("model_type", "hyb_type", "round", "accuracy", "accuracy_LL", "accuracy_UL",
                 "sensitivity", "specificity", "precision", "npv", "n",
                 "total_truths", "total_lies", "total_truth_guesses", "total_lie_guesses")

# part b: create matrix
results <-
  matrix(nrow = rounds,
         ncol = length(result_cols))

# part c: actually name columns in results marix
colnames(results) <- result_cols

# part d: convert to df (so multiple variables of different types can be stored)
results <- data.frame(results)

# -----------------------------------------------------------------------------
# STEP 2: start timer
start_time <- Sys.time()

# -----------------------------------------------------------------------------
# STEP 3: create rounds number of models, and store results each time

for (i in 1:rounds){
  
  # part a: partition data in 50-50 lgocv split (create index for test set)
  index_train <- 
    createDataPartition(y = stats_combo$stat_id,
                        p = 0.50,
                        times = 1,
                        list = FALSE)
  
  # part b: create testing and training data sets
  train_set <- stats_combo[index_train, ]
  test_set <- stats_combo[-index_train, ]
  
  
  # part c: use caret "train" function to train hybrid and non-hybrid logistic regression model
  model_hybrid_i <- 
  train(form = grd_truth ~ . - stat_id,
        data = train_set_ex,
        method = "glm",
        family = "binomial")

  model_non_hybrid_i <-
    train(form = grd_truth ~ . -stat_id -predict,
          data = train_set_ex,
          method = "glm",
          family = "binomial")
  
  
  # part d: make predictions, using both hybrid and non-hybrid models
  preds_hyb_i <-
    predict(object = model_hybrid_i,
            newdata = test_set,
            type = "raw")
  
  preds_non_hyb_i <-
    predict(object = model_non_hybrid_i,
            newdata = test_set,
            type = "raw")
  
  # part e: store model performance of both hybrid and non-hybrid model
  conf_hyb_i <-
    confusionMatrix(data = preds_hyb_i,
                    reference = test_set$grd_truth,
                    positive = "truth")
  
  conf_non_hyb_i <-
    confusionMatrix(data = preds_non_hyb_i,
                    reference = test_set$grd_truth,
                    positive = "truth")
  
  # part f: store hybrid and non-hybrod model results
  # model type
  results[2*i-1, 1] <- "logistic"
  results[2*i, 1] <- "logistic"
  # hybrid o non-hybrid
  results[2*i-1, 2] <- "hybrid"
  results[2*i, 2] <- "non_hybrid"
  # round
  results[2*i-1, 3] <- i
  results[2*i, 3] <- i
  # accuracy
  results[2*i-1, 4] <- conf_hyb_i$overall[1]
  results[2*i, 4] <- conf_non_hyb_i$overall[1]
  # accuracy LL
  results[2*i-1, 5] <- conf_hyb_i$overall[3]
   results[2*i, 5] <- conf_non_hyb_i$overall[3]
  # accuracy UL
  results[2*i-1, 6] <- conf_hyb_i$overall[4]
  results[2*i, 6] <- conf_non_hyb_i$overall[4]
  # sensitivity
  results[2*i-1, 7] <- conf_hyb_i$byClass[1]
  results[2*i, 7] <- conf_non_hyb_i$byClass[1]
  # specificity
  results[2*i-1, 8] <- conf_hyb_i$byClass[2]
  results[2*i, 8] <- conf_non_hyb_i$byClass[2]
  # precision
  results[2*i-1, 9] <- conf_hyb_i$byClass[3]
  results[2*i, 9] <- conf_non_hyb_i$byClass[3]
  # negative predictive value
  results[2*i-1, 10] <- conf_hyb_i$byClass[4]
  results[2*i, 10] <- conf_non_hyb_i$byClass[4]
  # sample size (total)
  results[2*i-1, 11] <- sum(conf_hyb_i$table)
  results[2*i, 11] <- sum(conf_non_hyb_i$table)
  # total truths
  results[2*i-1, 12] <- conf_hyb_i$table[3] + conf_hyb_i$table[4]
  results[2*i, 12] <- conf_non_hyb_i$table[3] + conf_hyb_i$table[4]
  # total lies
  results[2*i-1, 13] <- conf_hyb_i$table[1] + conf_hyb_i$table[2]
  results[2*i, 13] <- conf_non_hyb_i$table[1] + conf_non_hyb_i$table[2]
  # total truth guesses
  results[2*i-1, 14] <- conf_hyb_i$table[2] + conf_hyb_i$table[4]
  results[2*i, 14] <- conf_non_hyb_i$table[2] + conf_non_hyb_i$table[4]
  # total lie guesses
  results[2*i-1, 15] <- conf_hyb_i$table[1] + conf_hyb_i$table[3]
  results[2*i, 15] <- conf_non_hyb_i$table[1] + conf_non_hyb_i$table[3]
  
  # part g: print round and total elapsed time so far
  cumul_time <- difftime(Sys.time(), start_time, units = "mins")
  print(paste("round #", i, ": cumulative time ", round(cumul_time, 2), " mins",
              sep = ""))
  print("--------------------------------------")

}

```

## View Results (Tabular)

Directly below, I've displayed a raw tabular summary of the results. We have 20 rows,
which result from the 2 models (hybrid and non-hybrid) created in each of 10
rounds. One pattern that seems to stand out is that, although like before,
the hybrid model only has a slightly higher overall accuracy than the hybrid
model, this holds in every round. In the second chunk below, I have displayed this
directly. For each round, I print the overall accuracy rate of the hybrid model
next to the overall accuracy rate for the non-hybrid model. We see that in
literally every single round, the hybrid model performed better than the
non-hybrid model. This pattern of results seems unlikely to me, if the hybrid
and non-hybrid models truly were equivalent in performance. Indeed, when I 
conduct a simple sign test, which uses a binomial distribution to examine
how likely it is that one model should have a higher accuracy score than the other
model for 10 out of 10 rounds, under the null assumption that each model
should have a higher accuracy score an equal number of times, this pattern
diverges significanty from that null hypothesis, p = 0.002. Similarly,
when I conduct a Wilcoxon signed-rank test (which is non-parametric test which
tests whether there is sufficient evidence to reject the null hypothesis that two
*dependent* samples are drawn from the same population), this difference appears
to be significant, V = 55, p = 0.006. This is some evidence to suggest that
hybrid human-computer models can perform better than non-hybrid computer models,
which do not incorporate human judgments. (There may be some statistical artifact or
error that a more statistically savvy person may point out that I am failing to 
account for, but I cannot think of one. Perhaps there is some form of non-independence
that I have not accounted for.)


```{r}
# print results (long form)
results
  
```

```{r}
# get data in wide format (with accuracy for hybrid and non_hybrid models next to each other)
results_acc_wide <-
  results %>%
    select(hyb_type, round, accuracy) %>%
    spread(key = hyb_type,
           value = accuracy)

# print
results_acc_wide %>%
  mutate(winner = case_when(hybrid > non_hybrid ~ "hybrid",
                            hybrid < non_hybrid ~ "non-hybrid",
                            hybrid == non_hybrid ~ "tie"))

```

```{r}
# sign test, via binomial test
binom.test(x = 10,
           n = 10,
           p = 0.5)

```

```{r}
# conduct Wilcoxon signed-rank test 
wilcox.test(x = results_acc_wide$hybrid,
            y = results_acc_wide$non_hybrid,
            paired = TRUE,
            conf.int = 0.95)
```

## View Results (Graphically; Overall Accuracy)

When we visually examine the average overall accuracy rate of non-hybrid and hybrid 
models across 10 rounds, we see much the same pattern of results we saw when compared
these two model types only once in the example case with which we led. The hybrid 
model has a slightly higher level of overall accuracy than the non-hybrid model
(although, as we see the confidence intervals for the two models overall accuracy
rates overlap).

```{r}
# average n in test set for hybrid and non-hybrid models
total_hyb <- mean((results %>% filter(hyb_type == "hybrid"))$n)
total_non_hyb <- mean((results %>% filter(hyb_type == "non_hybrid"))$n)

# vector of denominators to use, when creating error bars
denoms_overall <- c(total_hyb,
                    total_non_hyb)

# visualize overall accuracy
results %>%
  group_by(hyb_type) %>%
  summarize(accuracy = mean(accuracy)) %>%
  ggplot(aes(x = hyb_type,
             y = accuracy)) +
  geom_point(size = 2,
             color = "#545EDF") +
  geom_errorbar(aes(ymin = accuracy - 1.96*sqrt(accuracy*(1-accuracy)/denoms_overall),
                     ymax = accuracy + 1.96*sqrt(accuracy*(1-accuracy)/denoms_overall)),
                color = "#545EDF",
                width = 0.05,
                size = 1) +
  geom_hline(yintercept = 0.5,
             linetype = "dashed",
             size = 0.5,
             color = "red") +
  scale_y_continuous(breaks = seq(from = 0.49, to = 0.70, by = 0.01),
                     limits = c(0.49, 0.70)) +
  scale_x_discrete(limits = c("non_hybrid", "hybrid")) +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(color = "grey",
                                          size = 0.25),
        panel.background = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5),
        axis.title.y = element_text(margin = 
                                      margin(t = 0, r = 10, b = 0, l = 0)),
        axis.title.x = element_text(margin = 
                                      margin(t = 10, r = 00, b = 0, l = 0))) +
  labs(title = "Accuracy by Model Type",
       x = "Model Type",
       y = "Overall Accuracy")
```

## View Results (Graphically; Sensitivity, Specificity, Precision, Negative Predictive Value)

We see a similar pattern of results when we examine the other four major performance
statistics average across 10 rounds. The hybrid model has a higher average level of
sensitivity, specificity, precision, and negative predictive value (although again
confidence intervals overlap).

```{r, fig.width=8, fig.height=8}
# average n in test set for hybrid and non-hybrid models
total_truths_hyb <- mean((results %>% filter(hyb_type == "hybrid"))$total_truths)
total_truths_non_hyb <- mean((results %>% filter(hyb_type == "non_hybrid"))$total_truths)
total_lies_hyb <- mean((results %>% filter(hyb_type == "hybrid"))$total_lies)
total_lies_non_hyb <- mean((results %>% filter(hyb_type == "non_hybrid"))$total_lies)
total_truth_guesses_hyb <- mean((results %>% filter(hyb_type == "hybrid"))$total_truth_guesses)
total_truth_guesses_non_hyb <- mean((results %>% filter(hyb_type == "non_hybrid"))$total_truth_guesses)
total_lie_guesses_hyb <- mean((results %>% filter(hyb_type == "hybrid"))$total_lie_guesses)
total_lie_guesses_non_hyb <- mean((results %>% filter(hyb_type == "non_hybrid"))$total_lie_guesses)

# vector of denominators to use, when creating error bars
denoms_perf <- c(total_truths_non_hyb,
                 total_truths_hyb,
                 total_lies_non_hyb,
                 total_lies_hyb,
                 total_truth_guesses_non_hyb,
                 total_truth_guesses_hyb,
                 total_lie_guesses_non_hyb,
                 total_lie_guesses_hyb)

# visualize results
results %>%
  select(hyb_type, round, sensitivity, specificity, precision, npv) %>%
  gather(key = "metric",
         value = "value",
         sensitivity, specificity, precision, npv) %>%
  group_by(hyb_type, metric) %>%
  summarize(value = mean(value)) %>%
  ungroup() %>%
  mutate(metric = factor(metric,
                            levels = c("sensitivity", "specificity", "precision", "npv"))) %>%
  ggplot(aes(x = hyb_type,
             y = value)) +
  geom_point(size = 2,
             color = "#545EDF") +
  geom_errorbar(aes(ymin = value - 1.96*sqrt(value*(1-value)/denoms_perf),
                     ymax = value + 1.96*sqrt(value*(1-value)/denoms_perf)),
                color = "#545EDF",
                width = 0.05,
                size = 1) +
  geom_hline(yintercept = 0.5,
             linetype = "dashed",
             size = 0.5,
             color = "red") +
  scale_y_continuous(breaks = seq(from = 0.50, to = 0.70, by = 0.05),
                     limits = c(0.49, 0.70)) +
  scale_x_discrete(limits = c("non_hybrid", "hybrid")) +
  facet_grid(metric ~ .) +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(color = "grey",
                                          size = 0.25),
        plot.background = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=1),
        axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5),
        axis.title.y = element_text(margin = 
                                      margin(t = 0, r = 10, b = 0, l = 0)),
        axis.title.x = element_text(margin = 
                                      margin(t = 10, r = 00, b = 0, l = 0))) +
  labs(title = "Metrics by Model Type",
       x = "Model Type",
       y = "Proportion")

```


# Save Results

For now, this concludes our analysis. As our last step, lets save our results. 

```{r save, eval=FALSE}
# rename results df, to be particular to this model type (for disambiguation later)
results_HYB_log <- results
model_hybrid_log <- model_hybrid
model_non_hybrid_log <- model_non_hybrid

# clear results variable
rm(results)

# save results in Rda file
save(results_HYB_log,
     # model_hybrid_log,
     # model_non_hybrid_log,
     file = "results_HYB_log.Rda")
```


# Citations

* Kuhn, M. (2008). Building predictive models in R using the caret package. 
Journal of Statistical Software, 28(5), 1-26.

* Kuhn, M. (2019). The caret Package. Retrieved from https://topepo.github.io/caret/index.html

# END